---
output:
  html_document: default
  pdf_document: default
---
bplist00?_WebMainResource?	
^WebResourceURL_WebResourceFrameName_WebResourceData_WebResourceMIMEType_WebResourceTextEncodingName_`https://moodle.helsinki.fi/pluginfile.php/4433747/mod_resource/content/1/Hands-On-Exercise1a.RmdPO\j<html><head></head><body><pre style="word-wrap: break-word; white-space: pre-wrap;">---
title: "Quantitative Research Skills 2022, Hands-On Exercise 1a"
subtitle: "Continuous and categorical variables, visualisation, data wrangling"

author: "Rong Guang"
date: "18092022"

output: 
  html_document:
    theme: flatly
    highlight: haddock
    toc: true
    toc_depth: 2
    number_section: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Hands-On Exercise 1a: Continuous and categorical variables, visualisation, data wrangling

This R Markdown sheet again includes a quite large set of R code chunks taken from the RHDS book
**R for Health Data Science** by Ewen Harrison and Riinu Pius (CRC Press, 2021), available online at https://argoshare.is.ed.ac.uk/healthyr_book/. 
The structure here follows the structure of the book (chapters, sections and subsections).

*******************************************************************************


**NOTE:** There is also the Hands-On Exercise **1b**. Both parts (1a &amp; 1b) will be required for **Assignment 1**. You can begin working with the 1b as soon as you have completed this **1a** part.


*******************************************************************************

Here, we will continue studying of the RHDS book, now from **Chapter 6**.

# RHDS Chapter 6: Working with continuous outcome variables

Start reading the book at

https://argoshare.is.ed.ac.uk/healthyr_book/chap06-h1.html

and while reading the book, work through the R code below and try to understand what you are doing.

(To save time, we have copy-pasted all the necessary R code here for you, so you can focus on activating them.)

## 6.1 Continuous data

## 6.2 The Question

## 6.3 Get and check the data

```{r}
# Load packages (and install the finalfit package if not yet installed)
library(tidyverse)
#install.packages("finalfit")
library(finalfit) # install.packages("finalfit")
library(gapminder)

# Create object gapdata from object gapminder
gapdata  <-  gapminder

glimpse(gapdata) # each variable as line, variable type, first values

missing_glimpse(gapdata) # missing data for each variable

ff_glimpse(gapdata) # summary statistics for each variable
ff_glimpse(gapdata, levels_cut = 142)
```

## 6.4 Plot the data

### 6.4.1 Histogram

```{r}
gapdata %>% 
  filter(year %in% c(2002, 2007)) %>%
  ggplot(aes(x = lifeExp)) +       # remember aes()
  geom_histogram(bins = 20) +      # histogram with 20 bars
  facet_grid(year ~ continent, scales = "free")     # optional: add scales="free"                                 
```

### 6.4.2 Quantile-quantile (Q-Q) plot

```{r}
  gapdata %>% 
  filter(year %in% c(2002, 2007)) %>%
  ggplot(aes(sample = lifeExp)) +      # Q-Q plot requires 'sample'
  geom_qq() +                          # defaults to normal distribution
  geom_qq_line(colour = "blue") +      # add the theoretical line
  facet_grid(year ~ continent)
```

### 6.4.3 Boxplot

```{r}
gapdata %>% 
  filter(year %in% c(2002, 2007)) %>%
  ggplot(aes(x = continent, y = lifeExp)) +
  geom_boxplot() +
  facet_wrap(~ year)
```

6.4.3 continued...

```{r}
gapdata %>%  
  filter(year %in% c(2002, 2007)) %>%
  ggplot(aes(x = factor(year), y = lifeExp)) +
  geom_boxplot(aes(fill = continent)) +     # add colour to boxplots
  geom_jitter(alpha = 0.4) +                # alpha = transparency
  facet_wrap(~ continent, ncol = 5) +       # spread by continent
  theme(legend.position = "none") +         # remove legend
  xlab("Year") +                            # label x-axis
  ylab("Life expectancy (years)") +         # label y-axis
  ggtitle(
    "Life expectancy by continent in 2002 v 2007") # add title
```

## 6.5 Compare the means of two groups

### 6.5.1 t-test

### 6.5.2 Two-sample t-tests

```{r}
ttest_data <- gapdata %>%                    # save as object ttest_data
  filter(year == 2007) %>%                   # 2007 only
  filter(continent %in% c("Asia", "Europe")) # Asia/Europe only

ttest_result <- ttest_data %>%               # example using pipe
  t.test(lifeExp ~ continent, data = .)      # note data = ., see below
ttest_result

ttest_result$p.value # Extracted element of result object

ttest_result$conf.int # Extracted element of result object
```

### 6.5.3 Paired t-tests

```{r}
paired_data <-  gapdata %>%             # save as object paired_data
  filter(year %in% c(2002, 2007)) %>%  # 2002 and 2007 only
  filter(continent == "Asia")          # Asia only

paired_data %>%      
  ggplot(aes(x = year, y = lifeExp, 
             group = country)) +       # for individual country lines
  geom_line()
```

6.5.3 continued...

```{r}
paired_table <-  paired_data %>%        # save object paired_data
  select(country, year, lifeExp) %>%   # select vars interest
  pivot_wider(names_from = year,       # put years in columns
              values_from = lifeExp) %>% 
  mutate(
    dlifeExp = `2007` - `2002`         # difference in means
  )
paired_table

# Mean of difference in years
paired_table %>% summarise( mean(dlifeExp) )

paired_data %>% 
  t.test(lifeExp ~ year, data = ., paired = TRUE)
```

### 6.5.4 What if I run the wrong test?

## 6.6 Compare the mean of one group: one sample t-tests

```{r}
# the tidy() function comes from the package broom - install it first!
library(broom) # install.packages("broom")
gapdata %>% 
  filter(year == 2007) %>%          # 2007 only
  group_by(continent) %>%           # split by continent
  do(                               # dplyr function
    t.test(.$lifeExp, mu = 77) %>%  # compare mean to 77 years 
      tidy()                        # tidy into tibble
  )

```

### 6.6.1 Interchangeability of t-tests

```{r}
# note that we're using dlifeExp
# so the differences we calculated above
t.test(paired_table$dlifeExp, mu = 0)
```

```{r}
paired_table <-  paired_data %>%        # save object paired_data
  select(country, year, lifeExp) %>%   # select vars interest
  pivot_wider(names_from = year,       # put years in columns
              values_from = lifeExp) %>% 
  mutate(
    dlifeExp = `2007` - `2002`)  # difference in means
paired_data
paired_table
  
t.test(paired_table$dlifeExp, mu = 0) 

t.test(paired_table$dlifeExp, mu = 0)
  
```
## 6.7 Compare the means of more than two groups

### 6.7.1 Plot the data

```{r}
gapdata %>% 
  filter(year == 2007) %>% 
  filter(continent %in% 
           c("Americas", "Europe", "Asia")) %>% 
  ggplot(aes(x = continent, y=lifeExp)) +
  geom_boxplot() +
  theme_grey()
```

### 6.7.2 ANOVA

```{r}
aov_data  <-  gapdata %>% 
  filter(year == 2007) %>% 
  filter(continent %in% c("Americas", "Europe", "Asia"))

fit = aov(lifeExp ~ continent, data = aov_data) 
summary(fit)
fit
tidy(fit)
```

6.7.2 continued...

```{r}
library(broom)
gapdata %>% 
  filter(year == 2007) %>% 
  filter(continent %in% c("Americas", "Europe", "Asia")) %>% 
  aov(lifeExp~continent, data = .) %>% 
  tidy()
```


### 6.7.3 Assumptions

```{r}
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)
# to try this, you must install the ggfortify package:
install.packages("ggfortify")
library(ggfortify) # install.packages("ggfortify")
autoplot(fit)
```

## 6.8 Multiple testing

### 6.8.1 Pairwise testing and multiple comparisons

```{r}
aov_data

pairwise.t.test(aov_data$lifeExp, aov_data$continent, 
                p.adjust.method = "bonferroni")

pairwise.t.test(aov_data$lifeExp, aov_data$continent, 
                p.adjust.method = "fdr")
```

## 6.9 Non-parametric tests

### 6.9.1 Transforming data

```{r}
africa2002  <-  gapdata %>%              # save as africa2002
  filter(year == 2002) %>%             # only 2002
  filter(continent == "Africa") %>%    # only Africa
  select(country, lifeExp) %>%         # only these variables
  mutate(
    lifeExp_log = log(lifeExp)         # log life expectancy
  )
head(africa2002)                       # inspect

africa2002
africa2002 %>% 
  # pivot lifeExp and lifeExp_log values to same column (for easy plotting):
  pivot_longer(contains("lifeExp")) %>% 
  ggplot(aes(x = value)) +             
  geom_histogram(bins = 15) +          # make histogram
  facet_wrap(~name, scales = "free")   # facet with axes free to vary

long <- pivot_longer(africa2002, contains("lifeExp")) # what does pivot_longer do?
long
```

### 6.9.2 Non-parametric test for comparing two groups

```{r}
africa_data  <-  gapdata %>%                          
  filter(year %in% c(1982, 2007)) %>%      # only 1982 and 2007
  filter(continent %in% c("Africa"))       # only Africa

p1  <-  africa_data %>%                      # save plot as p1
  ggplot(aes(x = lifeExp)) + 
  geom_histogram(bins = 15) +
  facet_wrap(~year)

p2  <-  africa_data %>%                      # save plot as p2
  ggplot(aes(sample = lifeExp)) +          # `sample` for Q-Q plot
  geom_qq() + 
  geom_qq_line(colour = "blue") + 
  facet_wrap(~year)

p3 <-  africa_data %>%                      # save plot as p3
  ggplot(aes(x = factor(year),             # try without factor(year) to
             y = lifeExp)) +               # see the difference
  geom_boxplot(aes(fill = factor(year))) + # colour boxplot
  geom_jitter(alpha = 0.4) +               # add data points
  theme(legend.position = "none")          # remove legend

#install.packages("patchwork")
library(patchwork)                         # great for combining plots
p1 / p1 | p2 / p2 | p3 

africa_data %>% 
  wilcox.test(lifeExp ~ year, data = .)
```

### 6.9.3 Non-parametric test for comparing more than two groups

```{r}
library(broom)
gapdata %>% 
  filter(year == 2007) %>% 
  filter(continent %in% c("Americas", "Europe", "Asia")) %>%  
  kruskal.test(lifeExp~continent, data = .)  %>%  
  tidy()
```

## 6.10 Finalfit approach

```{r}
dependent  <-  "year"
explanatory  <-  c("lifeExp", "pop", "gdpPercap")
africa_data  %>%          
  mutate(
    year = factor(year)  # change variable type
  )  %>%  
  summary_factorlist(dependent, explanatory,
                     cont = "median", p = TRUE) 

```

6.10 continued...

```{r}
africa_data
explanatory
dependent  <-  "year"
explanatory  <-  c("lifeExp", "pop", "gdpPercap")
africa_data %>%         
  mutate(
    year = factor(year)
  )  %>%  
  summary_factorlist(dependent, explanatory,
                     cont_nonpara =  c(1, 3),         # variable 1&amp;3 are non-parametric
                     cont_range = TRUE,               # lower and upper quartile
                     p = TRUE,                        # include hypothesis test
                     p_cont_para = "t.test",          # use t.test/aov for parametric
                     add_row_totals = TRUE,           # row totals
                     include_row_missing_col = FALSE, # missing values row totals
                     add_dependent_label = TRUE)      # dependent label 
```

## 6.11 Conclusions


Great job! Chapter 6 DONE. Next: Chapter 8. Continue reading and working...


*******************************************************************************


# RHDS Chapter 8: Working with categorical outcome variables

https://argoshare.is.ed.ac.uk/healthyr_book/chap08-h1.html

## 8.1 Factors

## 8.2 The Question

## 8.3 Get the data

```{r}
# Get the data from the boot package (that includes tools for bootstrapping methods):
meldata <-  boot::melanoma # Survival from Malignant Melanoma
meldata
```

## 8.4 Check the data

```{r}
library(tidyverse)
library(finalfit)
theme_set(theme_bw())
meldata  %>%  glimpse()
meldata
meldata  %>%  ff_glimpse()
```

## 8.5 Recode the data
```{r}
meldata <- meldata %>% 
  mutate(sex.factor = factor (sex) %>%    #an alternative
           fct_
         (
             "Female" = "0",
             "Male"   = "1"
           ) %>% 
           ff_label ("sex"),
         ulcer.factor = factor (ulcer) %>% 
           fct_recode (
             "Absent"  =  "0",
             "Present" =  "1"
           ) %>% 
           ff_label("Ulcerated tumor"),
         factor.status = factor(status) %>% 
           fct_recode(
             "Died melanoma"  =  "1",
             "Alive"          =  "2",
             "Died - Other causes"  =  "3") %>% 
           ff_label ("Status")
           )

meldata
ff_glimpse(meldata)
```


```{r}
meldata  <-  meldata  %>%  
  mutate(sex.factor =             # Make new variable  
           sex  %>%                 # from existing variable
           factor()  %>%            # convert to factor
           fct_recode(            # forcats function
             "Female" = "0",      # new on left, old on right
             "Male"   = "1")  %>%  
           ff_label("Sex"),       # Optional label for finalfit
         
         # same thing but more condensed code:
         ulcer.factor = factor(ulcer) %>% 
           fct_recode("Present" = "1",
                      "Absent"  = "0")  %>%  
           ff_label("Ulcerated tumour"),
         
         status.factor = factor(status)  %>%  
           fct_recode("Died melanoma"       = "1",
                      "Alive"               = "2",
                      "Died - other causes" = "3")  %>%  
           ff_label("Status"))

View(meldata) # take a look at the data!
```

## 8.6 Should I convert a continuous variable to a categorical variable?

```{r}
# Summary of age
meldata$age  %>%  
  summary() %>% 
  tidy()

exp_age1 <- meldata %>% 
  ggplot(aes(x = age)) +
  geom_histogram(bins = 20, fill = "deepskyblue", colour = "black", size = 0.3, alpha = 0.3)


exp_age2 <- meldata %>% 
  ggplot(aes(sample = age)) +
  geom_qq(colour = "black", size = 1, shape = 1) +
  geom_qq_line(colour = "blue")

exp_age3 <- meldata %>% 
  mutate (age.label = "age") %>% 
  ggplot (aes(x = age.label, y = age)) +
  geom_boxplot() +
  xlab("") +
  ylab("")
  
exp_age1|exp_age2|exp_age3
```

### 8.6.1 Equal intervals vs quantiles

```{r}


meldata$age.factor  %>% 
  summary()


head(meldata)
meldata <- meldata %>% 
  mutate (age.factor = 
            age %>% 
            cut (4))
head(meldata)
meldata$age.factor %>% 
  summary()
```

8.6.1 continued...

```{r}
install.packages("Hmisc")
library("Hmisc")
meldata  <-  meldata  %>%  
  mutate(
    age.factor = 
      age  %>% 
      Hmisc::cut2(g=4) # Note, cut2 comes from the Hmisc package
  )

meldata$age.factor  %>%  
  summary()

View(meldata) # take a look at the data!
```

8.6.1 continued...

```{r}
meldata <- meldata %>% 
  mutate(age.factor = 
           age %>% 
           cut (
             breaks = c(4,20,40,60,95), include.lowest = TRUE) %>% 
               fct_recode ("≤20"          = "[4, 20]",
                           "21 to 40"     = "(20, 40]",
                           "41 to 60"     = "(40, 60]",
                           ">60"         = "(60,95]") %>% 
               ff_label("Age(years)")
  )
             

head(meldata$age.factor)
View(meldata) # take a look at the data!
```

## 8.7 Plot the data

```{r}
head(meldata)
p1  <-  meldata  %>%  
  ggplot(aes(x = ulcer.factor, fill = status.factor)) + 
  geom_bar() + 
  theme(legend.position = "none")

p2  <-  meldata  %>%  
  ggplot(aes(x = ulcer.factor, fill = status.factor)) + 
  geom_bar(position = "fill") + 
  ylab("proportion")

library(patchwork)
p1 + p2
```

8.7 continued...

```{r}
p1  <-  meldata  %>%  
  ggplot(aes(x = ulcer.factor, fill = status.factor)) + 
  geom_bar(position = position_stack(reverse = TRUE)) + 
  theme(legend.position = "none")

p2  <-  meldata  %>%  
  ggplot(aes(x = ulcer.factor, fill = status.factor)) + 
  geom_bar(position = position_fill(reverse = TRUE)) + 
  ylab("proportion")

library(patchwork)
p1 + p2
```

8.7 continued...

```{r}
p1  <-  meldata  %>%  
  ggplot(aes(x = ulcer.factor, fill=status.factor)) + 
  geom_bar(position = position_stack(reverse = TRUE)) +
  facet_grid(sex.factor ~ age.factor) + 
  theme(legend.position = "none")

p2  <-  meldata  %>%  
  ggplot(aes(x = ulcer.factor, fill=status.factor)) + 
  geom_bar(position = position_fill(reverse = TRUE)) +
  facet_grid(sex.factor ~ age.factor)+ 
  theme(legend.position = "bottom") +
  ylab("proportion") # this line was missing in the book

p1 / p2
```


## 8.8 Group factor levels together - fct_collapse()

```{r}
head(meldata)

meldata <- meldata %>% 
  mutate(status_dss = 
           fct_collapse(
             status.factor, 
             "Alive" = c("Alive", "Died - other causes"))
          )
head(meldata)
View(meldata) # take a look at the data! 
```

## 8.9 Change the order of values within a factor - fct_relevel()

```{r}
# dss - disease specific survival
meldata$status_dss %>% levels()
meldata %>% count(status_dss)

meldata %>% 
  ggplot (aes(x = ulcer.factor, fill = status_dss)) +
  geom_bar (position = position_fill (reverse = TRUE)) +
  facet_grid (sex.factor ~ age.factor) +
  theme(legend.position = "bottom") +
  ylab("proportion")




meldata  <-  meldata  %>%  
  mutate(status_dss = status_dss  %>% 
           fct_relevel("Alive")
         )

meldata$status_dss  %>%  levels()
```

## 8.10 Summarising factors with finalfit

```{r}
library(finalfit)
meldata %>% 
  summary_factorlist(dependent   = "status_dss", 
                     explanatory = "ulcer.factor") 
```

8.10 continued...

```{r}
library(finalfit)
meldata  %>%  
  summary_factorlist(dependent = "status_dss", 
                     explanatory = 
                       c("ulcer.factor", "age.factor", 
                         "sex.factor", "thickness")
  ) %>% 
knitr::kable(align=c("l", "l", "r", "r", "r", "r"))
```

## 8.11 Pearson’s chi-squared and Fisher’s exact tests

### 8.11.1 Base R

```{r}
table(meldata$ulcer.factor, meldata$status_dss) 

# both give same result

with(meldata, table(ulcer.factor, status_dss))
```

8.11.1 continued...

```{r}
library(magrittr)

meldata %$%        # note $ sign here
  table(ulcer.factor, status_dss)

meldata %$%
  table(ulcer.factor, status_dss) %>%  
  prop.table(margin = 1)     # 1: row, 2: column etc.

meldata %$%        # note $ sign here
  table(ulcer.factor, status_dss)  %>%  
  chisq.test()

library(broom)
meldata %$%        # note $ sign here
  table(ulcer.factor, status_dss)  %>%  
  chisq.test()  %>%  
  tidy()
```

## 8.12 Fisher’s exact test

```{r}
meldata %$%        # note $ sign here
  table(age.factor, status_dss)  %>%  
  chisq.test()

meldata %$%        # note $ sign here
  table(age.factor, status_dss)  %>%  
  fisher.test() %>% 
  tidy()
```

## 8.13 Chi-squared / Fisher’s exact test using finalfit

```{r}
library(finalfit)
meldata  %>%  
  summary_factorlist(dependent   = "status_dss", 
                     explanatory = "ulcer.factor",
                     p = TRUE, add_dependent_label = TRUE) 

meldata  %>%  
  summary_factorlist(dependent = "status_dss", 
                     explanatory = 
                       c("ulcer.factor", "age.factor", 
                         "sex.factor", "thickness"),
                     p = TRUE)

t1 <- meldata  %>%  
  summary_factorlist(dependent = "status_dss", 
                     explanatory = 
                       c("ulcer.factor", "age.factor", 
                         "sex.factor", "thickness"),
                     p = TRUE,
                     p_cat = "fisher") 
t1
knitr::kable(t1, align=c("l", "l", "r", "r", "r", "r"))
```

```{r}
install.packages("knitr")
library("knitr")

explanatory = c("age.factor", "sex.factor", "obstruct.factor")
dependent = 'mort_5yr'
colon_s %>%
  summary_factorlist(dependent, explanatory, 
  p=TRUE, add_dependent_label=TRUE) -> t1
knitr::kable(t1, align=c("l", "l", "r", "r", "r"))
```

##8.14 Exercise
##8.14.1
```{r}
dependent = "status.factor"
explanatory = c("sex.factor", "ulcer.factor", "age.factor", "thickness")

meldata %>% 
  summary_factorlist (dependent, explanatory, cont = "median", cont_range = TRUE) %>% 
  knitr::kable()
  

```
##8.14.2
```{r}
head(meldata)
meldata %>%
  count(ulcer.factor, status.factor) %>%
  group_by(status.factor) %>%
  mutate(total = sum(n)) %>%
  mutate(percentage = round(100*n/total, 1)) %>% 
  mutate(count_perc = paste0(n, " (", percentage, ")")) %>% 
  select(-total, -n, -percentage) %>% 
  spread(status.factor, count_perc)

#change one line to "by age"
  meldata %>%
  count(age.factor, status.factor) %>%
  group_by(status.factor) %>%
  mutate(total = sum(n)) %>%
  mutate(percentage = round(100*n/total, 1))  %>% 
  mutate(count_perc = paste0(n, " (", percentage, ")")) %>% 
  select(-total, -n, -percentage) %>% 
  spread(status.factor, count_perc) %>% 
    knitr::kable(align = c("l", "r", "r", "r"))
  
#change one line to "by sex"
  meldata %>%
  count(sex.factor, status.factor) %>%
  group_by(status.factor) %>%
  mutate(total = sum(n)) %>%
  mutate(percentage = round(100*n/total, 1))  %>% 
  mutate(count_perc = paste0(n, " (", percentage, ")")) %>% 
  select(-total, -n, -percentage) %>% 
  spread(status.factor, count_perc) %>% 
    knitr::kable(align = c("l", "r", "r", "r"))

#"by ulcer using finalfit"
  dependent = "status.factor"
  explanatory = c("ulcer.factor", "age.factor" ,"sex.factor")
  
  meldata %>% 
    summary_factorlist (dependent, explanatory) %>% 
    knitr::kable(align = c("l", "r", "r", "r", "r", "r"))
```
(For some reason, the code chunk in 8.13, below the text "Further options can be
included" does not seem to work, so we will just skip it here.)

*******************************************************************************

**Good job!** 

Finally, continue with the RHDS Chapter 3 (below).

After that, you have actively read chapters 1, 2, 3, 4, 6, and 8 of the RHDS book.

We will leave Chapter 5 to you as an optional chapter. It is worth checking, if you want to fine-tune some of your graphs, for example for the **Assignment 5** (in the end of the course).

Have a good time with Chapter 3!

*******************************************************************************


# RHDS Chapter 3: Summarising data

Continue reading at 
https://argoshare.is.ed.ac.uk/healthyr_book/summarising-data.html 
and working with the R code chunks. Remember to write your own comments, too!

## 3.1 Get the data

```{r}
library(tidyverse)
gbd_full  <- read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/global_burden_disease_cause-year-sex-income.csv")

# Creating a single-year tibble for printing and simple examples:
gbd2017  <-  gbd_full  %>%  
  filter(year == 2017)

View(gbd2017)
head(gbd2017)
```

## 3.2 Plot the data

```{r}
gbd_pic <- gbd2017  %>%  
  # without the mutate(... = fct_relevel()) 
  # the panels get ordered alphabetically
  mutate(income = fct_relevel(income,
                              "Low",
                              "Lower-Middle",
                              "Upper-Middle",
                              "High")) %>%  
  # defining the variables using ggplot(aes(...)):
  ggplot(aes(x = sex, y = deaths_millions, fill = cause)) +
  # facets for the income groups:
  facet_wrap(~income, ncol = 4)  + 
  scale_fill_brewer(palette="Paired")

 # type of geom to be used: column (that's a type of barplot):
  gbd_pic1 <- gbd_pic + 
    geom_col() +
  # move the legend to the top of the plot (default is "right"):
  theme(legend.position = "top")
  
  gbd_pic2 <- gbd_pic + 
    geom_col(position = "fill") +
    theme(legend.position = "none")
  
  gbd_pic3 <- gbd_pic + 
    geom_col(position = "dodge") +
    theme(legend.position = "none")
library("patchwork")
gbd_pic1/gbd_pic2/gbd_pic3
  

```

## 3.3 Aggregating: group_by(), summarise()

```{r}
?summarise
gbd2017$deaths_millions  %>%  sum()

addsum <- gbd2017  %>%  
  group_by(sex) %>% 
  summarise(sum = sum(deaths_millions), mean = mean(deaths_millions), sd = sd(deaths_millions), n = n(), IQR = IQR(deaths_millions))
addsum

gbd2017  %>%  
  group_by(cause)  %>%  
  summarise(sum(deaths_millions))

gbd2017  %>%  
  group_by(cause, sex)  %>%  
  summarise(sum = sum(deaths_millions)) %>% 
  spread(sex, sum)
```

## 3.4 Add new columns: mutate()

```{r}
gbd2017  %>%  
  group_by(cause, sex)  %>%  
  summarise(deaths_per_group = sum(deaths_millions)) %>%  
  ungroup()  %>%  
  mutate(deaths_total = sum(deaths_per_group))
```

### 3.4.1 Percentages formatting: percent()

```{r}
# percent() function for formatting percentages come from library(scales)
install.packages("scales")
library(scales)
gbd2017_summarised  <-  gbd2017  %>%  
  group_by(cause, sex)  %>%  
  summarise(deaths_per_group = sum(deaths_millions))  %>%  
  ungroup()  %>%  
  mutate(deaths_total    = sum(deaths_per_group),
         deaths_relative = percent(deaths_per_group/deaths_total))
gbd2017_summarised

# using values from the first row as an example:
round(100*4.91/55.74, 1)  %>%  paste0("%")

gbd2017_summarised  %>%  
  mutate(deaths_relative = deaths_per_group/deaths_total)
```

## 3.5 summarise() vs mutate()

```{r}
head(gbd2017)

gbd_summarised  <-  gbd2017  %>%  
  mutate (sex= factor (sex)) %>%   #change sex variable to factor type
  mutate (sex = fct_relevel(sex, "Male", "Female")) %>%  #recorder sex variable
  group_by(cause, sex)  %>%  
  summarise(deaths_per_group = sum(deaths_millions))  %>%  
  arrange(sex)

gbd_summarised


gbd_summarised_sex  <-  gbd_summarised  %>%  
  group_by(sex)  %>%  
  summarise(deaths_per_sex = sum(deaths_per_group))

#another way to realize the above codes
gbd_summarised_sex

gbd_summarised_sex <- gbd_summarised %>% 
  group_by(sex) %>% 
  summarise (sum_sex = sum(deaths_per_group)) %>% 
  mutate(sex = sex %>% fct_relevel("Female", "Male")) %>% 
  arrange(sex)

gbd_summarised_sex 


```

3.5 continued...

```{r}
full_join(gbd_summarised, gbd_summarised_sex)

gbd_summarised  %>%  
  group_by(sex)  %>%  
  mutate(deaths_per_sex = sum(deaths_per_group))

gbd2017  %>%  
  group_by(cause, sex)  %>%  
  summarise(deaths_per_group = sum(deaths_millions))  %>%  
  group_by(sex)  %>%  
  mutate(deaths_per_sex  = sum(deaths_per_group),
         sex_cause_perc = percent(deaths_per_group/deaths_per_sex))  %>%  
  arrange(sex, deaths_per_group)
```

## 3.6 Common arithmetic functions - sum(), mean(), median(), etc.

```{r}
mynumbers  <-  c(1, 2, NA)
sum(mynumbers)

sum(mynumbers, na.rm = TRUE)
```

## 3.7 select() columns

```{r}
head(gbd_full)
gbd_2rows  <-  gbd_full  %>%  
  slice(1:2)

gbd_2rows

gbd_2rows  %>%  
  select(cause, deaths_millions)

gbd_2rows  %>%  
  select(cause, deaths = deaths_millions)

gbd_2rows  %>%  
  rename(deaths = deaths_millions)

gbd_2rows  %>%  
  select(year, sex, income, cause, deaths_millions)

gbd_2rows  %>%  
  select(year, sex, everything())

gbd_2rows  %>%  
  select(starts_with("deaths"))
```

## 3.8 Reshaping data - long vs wide format

```{r}
gbd_wide <-  read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/global_burden_disease_wide-format.csv")
gbd_long  <-  read_csv("https://raw.githubusercontent.com/KimmoVehkalahti/RHDS/master/data/global_burden_disease_cause-year-sex.csv")

gbd_wide
gbd_long
```

### 3.8.1 Pivot values from rows into columns (wider)

```{r}
gbd_long  %>%  
  pivot_wider(names_from = year, values_from = deaths_millions)

gbd_long  %>%  
  pivot_wider(names_from = sex, values_from = deaths_millions)  %>%  
  mutate(Male - Female)

gbd_long %>% 
  pivot_wider(names_from = c(sex, year), values_from = deaths_millions)
```

### 3.8.2 Pivot values from columns to rows (longer)

```{r}
gbd_wide
gbd_wide %>%  
  pivot_longer(matches("Female|Male"), 
               names_to = "sex_year", 
               values_to = "deaths_millions")  %>%  
  slice(1:6)
gbd_wide
gbd_wide  %>%  
  select(matches("Female|Male"))
```

### 3.8.3 separate() a column into multiple columns

```{r}
gbd_wide  %>%  
  # same pivot_longer as before
  pivot_longer(matches("Female|Male"), 
               names_to = "sex_year", 
               values_to = "deaths_millions")  %>%  
  separate(sex_year, into = c("sex", "year"), sep = "_", convert = TRUE)
```

## 3.9 arrange() rows

```{r}
gbd_long
gbd_long  %>%  
  arrange(deaths_millions)  %>%  
  # first 3 rows just for printing:
  slice(1:3) 

gbd_long  %>%  
  arrange(-deaths_millions)  %>%  
  slice(1:3)
gbd_long  %>%  
  arrange(desc(sex))  %>%  
  # printing rows 1, 2, 11, and 12
  slice(1,2, 11, 12)
```

### 3.9.1 Factor levels

```{r}
gbd_long
gbd_factored <-  gbd_long  %>%  
  mutate(cause = factor(cause))

gbd_factored$cause  %>%  levels()

gbd_factored  <-  gbd_factored  %>%  
  mutate(cause = cause %>% 
           fct_relevel("Injuries"))

gbd_factored$cause  %>%  levels()
```

Yes! That was all for Chapter 3. **GOOD JOB.**

*******************************************************************************

**Well done! That was active reading of Chapter 3.**

You have now completed the **Hands-On Exercise 1a**. *GOOD JOB!*

Please continue with the **Hands-on Exercise 1b** that will get you started with the **USHS data**.

(It is good to keep also the **1a** open, while working with the **1b**.)
</pre></body></html>_text/x-markdownUUTF-8