# Instructions for working with POR (SPSS Portable) files / KV 16 Oct 2022
#
# During the course, we have followed the RHDS book in working with
# *factors*, i.e., categorical variables that have two or more *levels*,
# for example 'gender' with the levels 1 = 'Male' and 2 = 'Female'.
#
# In addition, I have given examples of converting numeric variables
# of the USHS data to factors. The USHS data was originally given in the
# CSV (comma-separated values) form, which is a universal form, but it
# cannot handle the factors very well, at least if the variables are
# saved in the numeric form (as was the case with the USHS data).
#
# Now, I decided to use another way of reading data in R. FSD distributes
# their data sets also in the POR (SPSS Portable) form that preserves
# the textual information of the variables (in SPSS vocabulary: "variable
# labels" and "value labels").
#
# To work with the POR files, we will use the 'haven' package, that is also 
# part of tidyverse. You can download from Moodle one of the POR files (just
# click its name) or all of them (click "Download folder" and unzip to your
# working directory). Then you can start working in your R Markdown sheet
# as follows:
```{r}
library(tidyverse)
library(tidyr)
library(gapminder) # dataset
library(finalfit)
library(broom)
library(gtsummary)#install.packages("gtsummary")
library(sjlabelled)#install.packages("sjlabelled")
library(lubridate)#install.packages("lubridate")
library(sjlabelled)
library(psych)
library(Hmisc)
library(expss)#install.packages("expss")
library(patchwork)
library(ggfortify)
library(DT)
library(modelsummary)#install.packages("modelsummary")
```

```{r}
por1 <- "daF3077e.por"
por2 <- "daF3153e.por"
por3 <- "daF3172e.por"
por4 <- "daF3316e.por"
por5 <- "daF3345e.por"
por6 <- "daF3448e.por"
por7 <- "daF3494e.por"
por8 <- "daF3510.por"
```

por1 <- "daF3077e.por"
por2 <- "daF3153e.por"
por3 <- "daF3172e.por"
por4 <- "daF3316e.por"
por5 <- "daF3345e.por"
por6 <- "daF3448e.por"
por7 <- "daF3494e.por"
por8 <- "daF3510.por"
```{r}
library(haven) # part of tidyverse
library(dplyr) # part of tidyverse
```


# As an example, I will use the first data set from the above list:

# read the data in R:
```{r}
data1 <- read_por(por1)
```


# browse the data and check its structure:

```{r}
View(data1)
glimpse(data1) # note the "<dbl+lbl>": double & label! (haven speciality)

data1 %>% DT::datatable()
```



# summary of V5_3 (as a continuous variable - ERRONEOUS because of '9's!)
# remember to change '9's to NA's (as we did with the USHS data) if you
# use this variable in numeric form (in Factor Analysis, for example)!

```{r}
data1 %>%
  select(V5_3) %>%
  summary

data1 %>% ggplot(aes(x = V5_3)) +
  geom_boxplot()

data1 %>% select(V5_1:V5_4) %>% 
  pivot_longer(everything(), names_to = "items", values_to = "score") %>% 
  ggplot(aes(x = score)) +
  geom_boxplot() +
  facet_wrap(~items)
```


# summary of V5_3 as a factor:
# OBS: "as_factor" (from haven), NOT "as.factor" (from base R)!
data1 %>%
  select(V5_3) %>%
  as_factor() %>%
  summary

# ---------------------------------------------------
# NOTE!!
# it may be a good strategy to have TWO VERSIONS of the data:
#
#   1) numeric (see above) and 2) labelled (see below)!
#
# (with the latter one, it will be easier to understand the contents
# of the variables, while the numeric data is needed in Factor Analysis,
# for example)
#
# ---------------------------------------------------

# convert ALL the lbl's of the data to factors (dbl's will remain as dbl's):
# OBS: create a NEW DATA here with a new name (data1F)
```{r}
data1F <- as_factor(data1) # very powerful operation!
data1F
```


# that can be done only to a data set with "<dbl+lbl>" types! (haven speciality)

# browse the data in labelled form:
```{r}
View(data1F)
glimpse(data1F) # now each variable is either <fct> or <dbl> (usual R style)
```



# summary of V5_3 (now fixed as a factor):
```{r}
data1F %>%
  select(V5_3) %>%
  # as_factor() %>% # not needed anymore
  summary() # displays the 5 factor levels (labels) and their frequencies

```


# OBS: the levels do NOT have the numeric values anymore, ONLY the labels!

# so, if you need to recode the labels, you must do it following this:
# (you must use both the new and old labels, as there are no numbers)
```{r}
library(forcats)
data1F <- data1F %>% # save the changes back in the same data
  mutate(V5_3_reduced = V5_3 %>% # change the labels of the factor variable
    fct_recode("Agree" = "Strongly agree",
               "Agree" = "Somewhat agree", 
               "Disagree" = "Somewhat disagree", 
               "Disagree" = "Strongly disagree") %>%
    fct_rev() # this will also reverse the levels (sometimes handy!)
  ) # end of mutate
```


# the "Don't know, no answer" level was not changed, it is still there:

```{r}
data1F %>%
  select(V5_3, V5_3_reduced) %>%
  summary() # displays the 3 factor levels (labels) and their frequencies
```



# this is how you can access the variable label: (inherited via haven)
```{r}
attributes(data1F$V5_3)$label
var_lab(data1F$V5_3)
lapply(data1F, function(x)var_lab(x))
lapply(data1F, function(x)val_lab(x))
sapply(data1F[,-c(1:4)], function(x)table(x))
?pull
```


# The label is still quite long, so you may just copy+paste and shorten it,
# in case you need it in some plot titles etc. - for example (I just add a
# line break '\n' to have it on two lines in the plot, without shortening):]
```{r}
label_V5_3 <- "The EU shall enable and support more mobility of students \n and workers between the member states than existing today"
```


# let us create a bar plot of the factor variable V5_3:

```{r}
library(ggplot2)
p1 <- data1F %>%
  ggplot(aes(x = V5_3_reduced)) + 
    geom_bar(color = "blue", fill = "red") + 
    coord_flip() + 
    labs(title = label_V5_3)
p1
```




p1 # look at the plot

# Just for fun, let us save the plot in a PNG graphics file, with some options:
```{r}

ggsave("plot.png", plot = p1, units = "cm", width = 25, height = 15, dpi = 300)
```

# look at the Files pane of RStudio and click on the plot.png file to see it!
