---
title: "Quantitative Research Skills 2022, Assignment 4"
subtitle: "Logistic Regression"

author: "Write your name here"
date: "Write the date here"

output: 
  html_document:
    theme: flatly
    highlight: haddock
    toc: true
    toc_depth: 2
    number_section: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 4: Logistic Regression

Begin working with this when you have completed the **Hands-On Exercise 4**.

*********************************************

Show and report (step by step) the **Logistic Regression models** you created with the **USHS data**. What kind of conclusions do you make in each of the models? How about the odds-ratios of the explanatory variables?



# 1. Preparing 

## 1.1 Packages loading

```{r}
library(tidyverse)
library(tidyr)
library(gapminder) # dataset
library(finalfit)
library(broom)
library(gtsummary)#install.packages("gtsummary")
library(sjlabelled)#install.packages("sjlabelled")
library(lubridate)#install.packages("lubridate")
library(sjlabelled)
library(psych)
library(Hmisc)
library(expss)#install.packages("expss")
library(patchwork)
library(ggfortify)
library(DT)
```


## 1.2 Data loading
```{r}
USHSv3 <- read_csv("USHSv3")
USHSv3 %>% select (545:569) %>% DT::datatable()
```
## 1.3 Generating new variables

### 1.3.1 Generating variable about 1-year medical help seeking practice
```{r, eval = F}

# questions under k37 collected information about whether and how much medical
#help seeking behaviors (MHSB) the respondents had for the past one year for 
#varioustype of medical providers. Here I simplified them into whether they  
#have MHSB for any type of service. 
#select columns k37s

col_k37s <- USHSv3  %>% select (fsd_id, starts_with("k37"))
 
#How many types of medical provider he/she saw 1~5 times last year.

col_k37s$med_seek_face_few <- rowSums(col_k37s[-1] == 1, na.rm = T)
col_k37s <- col_k37s %>% mutate(med_seek_face_few = med_seek_face_few %>% ff_label("medical visits a few times"))

#How many types of medical provider he/she saw >5 times last year.
#a new variable
col_k37s$med_seek_face_many <- rowSums(col_k37s[-1] == 2, na.rm = T)
col_k37s <- col_k37s %>% mutate(med_seek_face_many = med_seek_face_many %>% ff_label("medical visits many times"))

get_labels(col_k37s)
```



```{r}
col_k37s <- col_k37s  %>% 
  rowwise() %>% 
  mutate(med_seek_face_few =sum(
    across (starts_with("k37")&contains("_2_")) == 1, na.rm = T)%>%
      ff_label("medical visits a few times"))

#How many types of medical provider he/she saw >5 times last year.
#a new variable
col_k37s <- col_k37s  %>% 
  rowwise() %>% 
  mutate(med_seek_face_many =sum(
    across (starts_with("k37")&contains("_2_")) == 2, na.rm = T) %>%
      ff_label("medical visits many times"))

#How many types of medical service he/she saw via phone last year.
#a new variable
col_k37s <- col_k37s  %>% 
  rowwise() %>% 
  mutate(med_seek_phone =sum(
    across (starts_with("k37")&contains("_3_")) == 3, na.rm = T))%>% 
  apply_labels(., med_seek_phone = "Medical phone call")

#How many types of medical service he/she saw via internet last year.
#a new variable
col_k37s <- col_k37s  %>% 
  rowwise() %>% 
  mutate(med_seek_online =sum(
    across (starts_with("k37")&contains("_4_")) == 4, na.rm = T))%>% 
    ungroup() %>% 
  apply_labels(., med_seek_online = "Medical help seeking on-line")


#If he/she sought any medical help during last year (yes/no)
#This is a combination of the four new variables above, if any of them has
#a value of 1, this new variable also gets a value of 1, or else it gets 0.

col_k37s <- col_k37s  %>% 
  mutate(med_seek_any = case_when(if_any(starts_with("med_seek"), ~.x != 0)~ "Yes",
                                  if_all(starts_with("med_seek"), ~.x == 0)~ "No") %>% 
           factor() %>% 
           ff_label("any medical help seeking during last year"))

          
#           sum(across (starts_with("med_seek")), na.rm = T)) %>% 
#  mutate(med_seek_any = if_else(med_seek_any == 0, 0, 1)) %>% 
#  ungroup()
#col_k37s %>% select(med_seek_any)  
#col_k37s <- col_k37s  %>% 
#  mutate(med_seek_any = med_seek_any %>%
#           factor() %>% 
#           fct_recode("Yes" = "1", "No" = "0") %>% 
#           ff_label("any medical help seeking during last year"))

USHSv2_col_k37s <- left_join(USHSv2, col_k37s)
USHSv2_col_k37s %>% select(starts_with("med_seek"))
write_csv(USHSv2_col_k37s, "USHSv2_col_k37s")

USHSv2_col_k37s %>% select(med_seek_face_many)
col_k37s %>% select(starts_with("med_seek")) 
```














*******************************************************

**GOOOOOD JOB!!**

Just knit this and SUBMIT the result (HTML) as your report of the **Assignment 4** in Moodle. 
